---
title: "p8105_hw2_alc2279"
author: "Annie Clark"
date: "September 29, 2018"
output: github_document
---
General questions - do we need to print everything? even when just load/clean data?
question 3 - a more elegant way to do proportion?
assumptions in question 3, part 2?
**make sure to name code chunks

#Problem 0
This “problem” focuses on structure of your submission, especially the use git 
and GitHub for reproducibility, R Projects to organize your work, R Markdown 
to write reproducible reports, relative paths to load data from local files, and 
reasonable naming structures for your files.

To that end:

* create a public GitHub repo + local R Project; we suggest naming this 
repo / directory p8105_hw2_YOURUNI (e.g.  p8105_hw2_ajg2202 for Jeff), but 
that’s not required
* create a single .Rmd file named p8105_hw2_YOURUNI.Rmd that renders to 
github_document
* create a subdirectory to store the local data files used in Problems 1 and 2,
and use relative paths to access these data files
* Your solutions to Problems 1, 2, and 3 should be implemented in your .Rmd file, 
and your git commit history should reflect the process you used to solve these 
problems.

For this Problem, we will assess adherence to the instructions above regarding 
repo structure, git commit history, and whether we are able to knit your .Rmd 
to ensure that your work is reproducible. Adherence to appropriate styling and 
clarity of code will be assessed in Problems 1+.

#Problem 1
##Part 1: Read and clean data
This problem focuses on NYC Transit data; in particular, this CSV file contains 
information related to each entrance and exit for each subway station in NYC.

Read and clean the data; retain line, station, name, station latitude / longitude, 
routes served, entry, vending, entrance type, and ADA compliance. Convert the entry 
variable from character (YES vs NO) to a logical variable (the ifelse or recode 
function may be useful).

```{r}
library(tidyverse)

subway_data = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(entry = entry == "YES")
```

##Part 2: Describe dataset
###Prompt:
Write a short paragraph about this dataset – explain briefly what variables the 
dataset contains, describe your data cleaning steps so far, and give the dimension 
(rows x columns) of the resulting dataset. Are these data tidy?

###Response
The dataset, subway_data, includes charcteristics of every subway entrance or exit;
more specifically, it includes data about what lines and routes can be accessed by each 
entrance/exit (line and route1:route11), its location, (station_latitude, station_longitude), 
and some important characteristics, like if entry is allowed as opposed to just an exit 
(entry), if there is vending available to buy tickets (vending), how the entrance/exit is
accessible (i.e. by stair, elevator, etc.) (entrance_type), and lastly if the entrance/exit
is approved by the ADA (ada). In the original dataset, there was additional information
included, like whether there is staffing, if crossovers are available (to a different
line) and several variables about the location and cross streets of the entrance. These
additional variables are not important to us for this exercise. 

So far, I have paired down the data to the relevant variables, standardized the names
(i.e. made them lower case and removed spaces/special characters), and converted
the entry variable from character to logical, which will be useful in the analysis. 

So far, this tabel that consists of `r nrow(subway_data)` rows and  `r ncol(subway_data)` 
columns is not yet considered to be "tidy." The 11 columns that contain route information
need to be gathered into one variable (and additionally cleaned up), the vending variable
should be changed to logical, the entrance_type variable could probably be changed to 
factor, and I would want to play around with cleaning up station_name and line (i.e. 
standardizing the case and possible the names, i.e. avenue vs. ave). 

##Part 3: Answer some questions

* How many distinct stations are there? Note that stations are identified both 
by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct 
function may be useful here.

```{r}
subway_data %>% 
  distinct(.data = ., station_name, line) %>% 
  nrow(.)
```

* How many stations are ADA compliant?

```{r}
subway_data %>%
  filter(ada == TRUE) %>%
  distinct(.data = ., station_name, line) %>% 
  nrow(.)
```

* What proportion of station entrances / exits without vending allow entrance?

```{r}
subway_data %>% 
  filter(vending == "NO") %>%
  pull(entry) %>% 
  mean()
```


##Part 4: Another exericse
Reformat data so that route number and route name are distinct variables. 
```{r}
subway_data =  
  gather(subway_data, key = route_num, value = route_name, route1:route11) %>% 
  separate(route_num, into = c("route", "route_num"), sep = 5) %>%
  select(line:ada, route_num, route_name)
```

How many distinct stations serve the A train?

```{r}
subway_data %>% 
  filter(route_name == "A") %>% 
  distinct(.data = ., station_name) %>% 
  nrow(.)
```

Of the stations that serve the A train, how many are ADA compliant?

```{r}
subway_data %>% 
  filter(route_name == "A", ada == TRUE) %>% 
  distinct(.data = ., station_name) %>% 
  nrow(.)
```

#Problem 2

##Part 1: Read and clean data
This problem uses the Mr. Trash Wheel dataset, available as an Excel file on the 
course website. Please use the  HealthyHarborWaterWheelTotals2017-9-26.xlsx version.

Read and clean the Mr. Trash Wheel sheet:

* specify the sheet in the Excel file and to omit columns containing notes (using 
the range argument and cell_cols() function)
* use reasonable variable names
* omit rows that do not include dumpster-specific data
* round the number of sports balls to the nearest integer and converts the result 
to an integer variable (using as.integer)

```{r}
library(readxl)

water_wheel_data = read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
                              sheet = "Mr. Trash Wheel",
                              range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster),
         month != "Grand Total") %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
```

Read and clean precipitation data for 2016 and 2017. For each, omit rows without 
precipitation data and add a variable year. 

```{r}
precipitation_2016_data = read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
                              sheet = "2016 Precipitation",
                              range = "A2:B14") %>%
  janitor::clean_names() %>% 
  rename(precipitation_in = total) %>% 
  filter(!is.na(precipitation_in)) %>% 
  mutate(year = "2016")

precipitation_2017_data = read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
                              sheet = "2017 Precipitation",
                              range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  rename(precipitation_in = total) %>% 
  filter(!is.na(precipitation_in)) %>% 
  mutate(year = "2017")
```

Next, combine datasets and convert month to a character variable (the variable 
month.name is built into R and should be useful).

```{r}
precipitation_data = bind_rows(precipitation_2016_data, precipitation_2017_data) %>% 
  mutate(month = month.name[month])
```

##Part 2: Describe dataset
Write a paragraph about these data; you are encouraged to use inline R. Be sure to 
note:
* the number of observations in both resulting datasets
* give examples of key variables
For available data, what was the total precipitation in 2017? What was the median 
number of sports balls in a dumpster in 2016?

```{r}
sum(precipitation_2017_data$precipitation_in)

water_wheel_data %>% filter(year == "2016") %>% pull (sports_balls) %>% median()
```


The first dataset created includes data about the volume/weight and type of 
trash picked up by the the water wheels (water_wheel_data) by date going 
back to May 2014. The data in its current state is not tidy (i.e. the year
variable has missing data, which can easily be cleaned). This data contains
'r nrow(water_wheel_data)' rows and 'r ncol(water_wheel_data)' columns. 

The second dataset includes precipitation data for 2016 and 2017. It was 
created by combining the 2016 dataset 'r nrow(precipitation_2016_data)' rows
and 'r ncol(precipitation_2016_data) columns) and the 2017 dataset ('r
nrow(precipitation_2017_data)' rows and 'r ncol(precipitation_2017_data
columns). The resulting data is 'r nrow(precipitation_data)' rows and 'r
ncol(precipitation_data) columns

#Problem 3

##Part 1: Read and clean data
This problem uses the BRFSS data. DO NOT include this dataset in your local data 
directory; instead, load the data from the  p8105.datasets package.

For this question:

* format the data to use appropriate variable names
* focus on the “Overall Health” topic
* exclude variables for class, topic, question, sample size, and everything from 
lower confidence limit to GeoLocation
* structure data so that responses (excellent to poor) are variables taking the 
value of Data_value
* create a new variable showing the proportion of responses that were “Excellent” 
or “Very Good”

```{r}
devtools::install_github("p8105/p8105.datasets")

library(p8105.datasets)

brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr, county = locationdesc) %>% 
  filter(topic == "Overall Health") %>% 
  select(-(class:question), -sample_size, -(confidence_limit_low:geo_location)) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  select(year:county, excellent, very_good, good, fair, poor) %>% 
  mutate(high_response = (excellent + very_good)) #adding excellent and high proportions
```

##Part 2: Answer some questions
Using this dataset, do or answer the following: 

* How many unique locations are included in the dataset? 

```{r}
#number of unique locations (county/state)
brfss_data %>% 
  distinct(.data = ., county) %>% 
  nrow()
```

* Is every state represented? 

```{r}
#number of states
brfss_data %>% 
  distinct(.data = ., state) %>% 
  nrow()

#view states - identify the extra one
distinct(brfss_data, state)

## DC is included
```

There are 51 states represented, including DC. 

* What state is observed the most?

```{r}
brfss_data %>% 
  count(state) %>% 
  arrange(desc(n)) %>% 
  pull(state) %>% 
  head(1)
```

* In 2002, what is the median of the “Excellent” response value?

```{r}
brfss_data %>%
  filter(year == 2002) %>% 
  pull(excellent) %>% 
  median(na.rm = TRUE)
```

* Make a histogram of “Excellent” response values in the year 2002.

```{r}
ggplot(brfss_data %>% filter(year == 2002), aes(x = excellent)) + 
  geom_histogram()
```

* Make a scatterplot showing the proportion of “Excellent” response values in New 
York County and Queens County (both in NY State) in each year from 2002 to 2010.

```{r}
ggplot(brfss_data %>% filter(county %in% c("NY - Queens County", "NY - New York County")) , 
  aes(x = year, y = excellent)) + 
  geom_point(aes(color = county))
```

